---
title: "Functions-guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions-guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PaleoPackage)
```

```{r include=FALSE}
devtools::load_all(".")
```

## Introduction to PaleoPackage

The PaleoPackage is designed to simplify the estimation and analysis of various anatomical and ecological parameters of extinct organisms, such as body mass, length, and speed. This package provides tools to work with fossil data and offers robust methods based on volumetric scaling, biomechanical principles, and anatomical measurements.
In this guide, we describe how to use each main function of the package, including detailed examples to help you get started with analyzing and interpreting fossil data effectively.

## Getting started 

Below is a quick demostration of some of the core functionalities of PaleoPackage. We will walk through how to:

- Estimate the body mass of an organism using different methods, such as `allometric_mass()` function.

- Calculate body length based on anatomical measurements.

- Predict running speed based on leg length and body mass.

Each example is designed to give you a practical understanding of how to use the functions effectively with fossil data. Let's dive in!

### Estimating Body Mass

The `allometric_mass()` function allows you to estimate the **body mass** of an organism based on its femur measurements. Here is a quick example:

```{r example}
# Example data
reference_weight <- 8800   # Weight of the reference organism in kilograms
reference_femur <- 136.5   # Femur length of the reference organism in centimeters
estimated_femur <- 140     # Femur length of the organism to estimate in centimeters

# Estimate body mass
mass <- allometric_mass(reference_weight, reference_femur, estimated_femur)
print(mass)
```

### Calculating the Cursoriality for Bipedal Animals

The `bipedal_cursoriality()` function calculates the **adjusted cadence** (steps per second) and **adjusted stride ratio** for bipedal aniamls based on their **body mass** and **leg length ratio**. These parameters help assess the locomotor adaptations and cursorial potential of bipedal animals across various weight ranges and anatomical proportions. Here's an example:

```{r}
# Example data
weight_kg <- 500          # Body mass of the animal in kilograms
leg_ratio <- 1.5          # Ratio of (Tibia + Metatarsal III) / Femur

# Calculate adjusted cursorial parameters
cursorial_params <- bipedal_cursoriality(weight_kg, leg_ratio)
print(cursorial_params)
```

### Calculating the Velocity for Bipedal Animals

The `bipedal_velocity()` function calculates the **velocity** (in kilometers per hour) of bipedal animals based on their **leg length**, **cadence**, and **stride ratio**. This function is particularly useful for estimating the locomotor capabilities of theropods and other bipedal organisms in paleontological research.

```{r}
# Example data
leg_length_mm <- 3167         # Length of the leg in millimeters (Femur + Tibia + Metatarsal III)
cadence <- 1.395             # Cadence of the animal in steps per second (Hz)
stride_ratio <- 1.424        # Stride ratio: proportion between stride length and leg length

# Calculate velocity
velocity_kmh <- bipedal_velocity(leg_length_mm, cadence, stride_ratio)
print(velocity_kmh)
```

